Off the beaten path: \\ Gaia reveals GD-1 stars perturbed off of the main stream track

### 2018-04-29 paper/gd1.tex
Figure placeholder

@@

Its long length --- and therefore its implied old dynamical age --- combined
with its particular orbit also make it a prime stream to search for interactions
with dark matter subhalos:
because of its large pericentric distance and retrograde orbit with respect to
the Galactic bar, density variations in the stream are not expected from either
interactions with giant molecular clouds (\citealt{Amorisco:2016}) or resonant
encounters with the bar (\citealt{Pearson:2017}).
By estimating a stream age of $\approx 4~\textrm{Gyr}$ assuming a progenitor
mass of $10^5~\msun$, and by taking a subhalo mass function and density from
XXX, \citet{Erkal:2016} predict that GD-1 could have up to one significant, wide
($\approx 5^\circ$--$7^\circ$) gap caused by an interaction with a
$10^6$--$10^7~\msun$ subhalo.

@@

We use astrometric data from the \gaia\ mission (\citealt{Prusti:2016}), data
release 2 (\citealt{Gaia-Collaboration:2018, Lindegren:2018}), and photometry
from the \pans\ survey, data release 1 (\citealt{Chambers:2016}).

We retrieve \gaia\ data along the previously-identified track of the GD-1 stream
from the \gaia\ science archive\footnote{\url{https://gea.esac.esa.int/}} by
selecting all sources with small parallax, $\varpi < 1~\textrm{mas}$, and use a
set of spherical polygons to select batches of stars that have small latitude in
the heliocentric GD-1 coordinate system, $(\phi_1, \phi_2)$, defined in
\cite{Koposov:2010}.\footnote{This transformation is implemented using the
\package{Astropy} (\citealt{astropy}) coordinate transformation framework in the
\package{Gala} \texttt{Python} package (\citealt{gala}).}
We convert the sky coordinates and proper motions of the parallax-selected stars
to the GD-1 stream coordinate system, and correct the proper motions for solar
reflex motion by assuming that stars at a given stream longitude, $\phi_1$, have
a distance given by the linear relationship $d(\phi_1) = (0.05 \, \phi_1 +
10)~\textrm{kpc}$.
For the solar velocity in a Galactocentric rest frame, we use $\bs{v}_\odot =
(11.1, 232.24, 7.25)~\kms$ (\citealt{Schonrich:2010, Bovy:2015}).
\figurename~\ref{fig:XX} (top right) shows the distribution of stars with
$|\phi_2| < XX^\circ$ in solar-reflex-corrected proper motion components in the
GD-1 coordinate system, $\mu_{\phi_1, \odot}$ (which includes the $\cos{\phi_2}$
term) and $\mu_{\phi_2, \odot}$.
GD-1 member stars are visible as the over-density at approximately
$(\mu_{\phi_1, \odot}, \mu_{\phi_2, \odot}) \approx (XX, YY)~\masyr$.

As a first selection of GD-1 candidate member stars, we select stars with proper
motions $XX < \mu_{\phi_1, \odot} < XX~\masyr$ and $XX < \mu_{\phi_2, \odot} <
XX~\masyr$, as visualized by the rectangle in the top right panel of
\figurename~\ref{fig:XX}.
\figurename~\ref{fig:XX} (top left) shows all stars the pass the selection on
proper motion, plotted in the GD-1 coordinate system.
From kinematic selection alone, the stream is identifiable as the over-density
of stars around $\phi_2 = 0$ between longitudes $XX^\circ \lesssim \phi_1
\lesssim XX^\circ$.

To improve the contrast of the stream over the background, we next cross-match
the sample to the \pans\ photometric catalog and use the photometry to further
clean the sample.
\figurename~\ref{fig:XX} (bottom right) shows the distribution of
proper-motion-selected candidate stream stars in a de-reddened \pans\
color-magnitude diagram.
The over-plotted isochrone is from the MESA Isochrones & Stellar Tracks (MIST;
\citealt{XX}) and represents a $12~\textrm{Gyr}$ old population with $\feh =
-1.XX$ at a distance of $XX~\kpc$.

- Color-magnitude filtering

>>>>

\newcommand{\todo}[1]{{\color{red} TODO: #1}}

@@

Its long length --- and therefore its implied old dynamical age --- combined
with its particular orbit also make it a prime stream to search for interactions
with dark matter subhalos:
because of its large pericentric distance and retrograde orbit with respect to
the Galactic bar, density variations in the stream are not expected from either
interactions with giant molecular clouds (\citealt{Amorisco:2016}) or resonant
encounters with the bar (\citealt{Pearson:2017}).
By estimating a stream age of $\approx 4~\textrm{Gyr}$ assuming a progenitor
mass of $10^5~\msun$, and by taking a subhalo mass function and density from
XXX, \citet{Erkal:2016} predict that GD-1 could have up to one significant, wide
($\approx 5$--$7^\circ$) gap caused by an interaction with a
$10^6$--$10^7~\msun$ subhalo.

@@

We use astrometric data from the \gaia\ mission (\citealt{Prusti:2016}), data
release 2 (\citealt{Gaia-Collaboration:2018, Lindegren:2018}), and photometry
from the \pans\ survey, data release 1 (\citealt{Chambers:2016}) to select
high-confidence members of the GD-1 stream.

\subselection{Stream member selection}

We retrieve \gaia\ data along the previously-identified track of the GD-1 stream
from the \gaia\ science archive\footnote{\url{https://gea.esac.esa.int/}} by
selecting all sources with small parallax, $\varpi < 1~\textrm{mas}$, and use a
set of spherical polygons to select batches of stars that have small latitude in
the heliocentric GD-1 coordinate system, $(\phi_1, \phi_2)$, defined in
\cite{Koposov:2010}.\footnote{This transformation is implemented using the
\package{Astropy} (\citealt{astropy}) coordinate transformation framework in the
\package{Gala} \texttt{Python} package (\citealt{gala}).}
We convert the sky coordinates and proper motions of the parallax-selected stars
to the GD-1 stream coordinate system, and correct the proper motions for solar
reflex motion by assuming that stars at a given stream longitude, $\phi_1$, have
a distance given by the linear relationship $d(\phi_1) = (0.05 \, \phi_1 +
10)~\textrm{kpc}$.
For the solar velocity in a Galactocentric rest frame, we use $\bs{v}_\odot =
(11.1, 232.24, 7.25)~\kms$ (\citealt{Schonrich:2010, Bovy:2015}).
\figurename~\ref{fig:selection} (top right) shows the distribution of stars with
$|\phi_2| < XX^\circ$ in solar-reflex-corrected proper motion components in the
GD-1 coordinate system, $\mu_{\phi_1, \odot}$ (which includes the $\cos{\phi_2}$
term) and $\mu_{\phi_2, \odot}$.
GD-1 member stars are visible as the over-density at approximately
$(\mu_{\phi_1, \odot}, \mu_{\phi_2, \odot}) \approx (XX, YY)~\masyr$.

As a first selection of GD-1 candidate member stars, we select stars with proper
motions $XX < \mu_{\phi_1, \odot} < XX~\masyr$ and $XX < \mu_{\phi_2, \odot} <
XX~\masyr$, as visualized by the rectangle in the top right panel of
\figurename~\ref{fig:selection}.
\figurename~\ref{fig:selection} (top left) shows all stars the pass the
selection on proper motion, plotted in the GD-1 coordinate system.
From kinematic selection alone, the stream is identifiable as the over-density
of stars around $\phi_2 = 0$ between longitudes $XX^\circ \lesssim \phi_1
\lesssim XX^\circ$.

\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{stream-selection.pdf}
\end{center}
\caption{
    \todo{AB}
\label{fig:selection}
}
\end{figure}

To improve the contrast of the stream over the background, we next cross-match
the sample to the \pans\ photometric catalog and use the photometry to further
clean the sample.
\figurename~\ref{fig:selection} (bottom right) shows the distribution of
proper-motion-selected candidate stream stars in a de-reddened \pans\
color-magnitude diagram.
\todo{Did you also cut on $phi_2$?}.
The over-plotted isochrone is from the MESA Isochrones & Stellar Tracks (MIST;
\citealt{XX}) and represents a $12~\textrm{Gyr}$ old population with $\feh =
-1.XX$ at a distance of $XX~\kpc$.
We use this isochrone to motivate a polygonal selection in de-reddened $g-i$
color and apparent $g$-band magnitude, as visualized by the \todo{shaded} region
in the bottom right panel of \figurename~\ref{fig:selection}.

\figurename~\ref{fig:selection} (bottom left) shows the final sample of GD-1
stream member candidates after selecting on both proper motions and photometry.
\todo{Comment on how this is the bomb ass diggity and the clearest view of the
stream so far}

\subselection{Stream member selection}
